Realizing temporal graphs from fastest travel times

### 2023-04-26 MFCS23/MFCS23-shortversion.tex
MFCS

The \emph{graph realization} problem with respect to a graph property $\mathcal{P}$ is to find a graph that satisfies property $\mathcal{P}$, or to decide that no such graph exists. 
The motivation for graph realization problems stems both from ``verification'' applications, 
where we attempt to reconstruct an existing network of unknown structure based on the outcomes of experimental measurements, 
and from network design applications in engineering, where the goal is to design network topologies having a desired property~\cite{augustine2022distributed,grotschel1995design}.
Analyzing the computational complexity of the graph realization problems for various natural and fundamental graph properties $\mathcal{P}$ requires a deep understanding of these properties.
Among the most well studied such parameters for graph realization 
are constraints on the distances between vertices~\cite{barNoy2022GraphRealization,barNoy2021composed,hakimi1965distance,chung2001distance,bixby1988almost,culberson1989fast}, 
on the vertex degrees~\cite{GolovachM17,gomory1961multi,hakimi1962realizability,Bar-NoyCPR20,erdos1960graphs}, 
%%%%barNoy2021selected,cai2000solution
on the eccentricities~\cite{barNoy2020efficiently,hell2009linear,behzad1976eccentric,lesniak1975eccentric}, and on connectivity~\cite{fulkerson1960zero,frank1992augmenting,chen1966realization,frank1994connectivity,frank1970connectivity,gomory1961multi}, among others.

In the simplest version of a graph realization problem with respect to vertex distances, 
we are given a symmetric $n \times n$ matrix $D$ and we are looking for an $n$-vertex undirected and unweighted graph $G$ such that $D_{i,j}$ equals the distance between vertices $v_i$ and $v_j$ in~$G$. This problem can be trivially solved in polynomial time in two steps~\cite{hakimi1965distance}: First, we build the graph $G=(V,E)$ such that $v_i v_j \in E$ if and only if $D_{i,j}=1$. Second, from this graph $G$ we compute the matrix $D_G$ which captures the shortest distances for all pairs of vertices. If $D_G = D$ then $G$ is the desired graph, otherwise there is no graph having $D$ as its distance matrix. 
Non-trivial variations this problem have been extensively studied, such as for weighted graphs~\cite{hakimi1965distance,Patrinos-Hakimi-72}, as well as for cases where the realizing graph has to belong to a specific graph family~\cite{hakimi1965distance, barNoy2021composed}. Other variations of the problem include the cases where every entry of the input matrix $D$ may contain a range of consecutive permissible values~\cite{barNoy2021composed,Rubei16,Tamura93}, or even an arbitrary set of acceptable values~\cite{barNoy2022GraphRealization} for the distance between the corresponding two vertices.

@@

In an arbitrary temporal graph $(G,\lambda)$ there can be pairs of vertices $u,v$ where there is no temporal path at all from $u$ to $v$, even if the underlying graph $G$ is undirected and connected. 
In order to focus on the smallest duration of temporal graphs, rather than on temporal reachability between vertices, we consider in this paper only \emph{periodic} temporal graphs, i.e.~on temporal graphs in which the availability times of every edge are periodic (all edges having the same period $\Delta$)~\cite{casteigts2012time}. 
Apart from the theoretical interest of periodic temporal graphs, which have not been well studied and understood in the literature, they capture naturally many application areas too, such as transportation or satellite networks, where entities move according to repeating schedules or time tables. For example, a fastest temporal path from vertex $u$ to vertex $v$ in the temporal graph defined by a transportation (resp.~satellite) network represents the shortest travel time (resp.~the shortest communication delay) from $u$ to $v$. 
As it turns out, the periodic temporal graph realization problem with respect to a given $n \times n$ matrix $D$ of the fastest duration times has a very different computational complexity behavior than the classic graph realization problem with respect to shortest path distances in static graphs.

@@

\subparagraph{Our contribution.}
%In this work, 
We initiate the study of naturally motivated graph realization problems in the temporal setting. To this end, we introduce and investigate the following problem, capturing the setting described above:

@@

We start with showing NP-hardness of the problem (\cref{thm:NPhardness}), even if $\Delta$ is a small constant. To establish a baseline for tractability, we show that \deltaExact\ is polynomial-time solvable if the underlying graph is a tree (\cref{thm:deltaExact-PolyTimeTrees}).

@@

We obtain the following results. We show that \deltaExact\ is W[1]-hard when parameterized by the feedback vertex number of the underlying graph (\cref{thm:W1wrtFVS}). 
To this end, we first give a reduction from \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval} to a variant of \deltaExact\ where the period $\Delta$ is infinite, that is, when the labeling is non-periodic. We use a special gadget (the ``infinity'' gadget) which allows us to transfer the result to a finite period $\Delta$. The latter construction is independent from the particular reduction we use, and can hence be treated as a reduction from the non-periodic to the periodic setting.

We complement this result by showing that the problem is fixed-parameter tractable (FPT) with respect to the feedback edge number $k$ of the underlying graph (\cref{thm:FPTwrtFES}). 
Our FPT algorithm works as follows on a high level. 
First we distinguish $O(k)$ vertices which we call ``interesting vertices''. 
Then, we guess the fastest temporal paths for each pair of these interesting vertices; as we prove, the number of choices we have for all these guesses is upper bounded by a function of $k$. 
Then we also make several further guesses (again using a bounded number of choices), which altogether leads us to specify a small (i.e.~bounded by a function of $k$) number of different configurations for the fastest paths between \emph{all pairs} of vertices. For each of these configurations, we must then make sure that the labels of our solution will not allow any other temporal path from a vertex $v_i$ to a vertex $v_j$ have a \emph{strictly smaller} duration than $D_{i,j}$.
This naturally leads us to build one Integer Linear Program (ILP) for each of these configurations. We manage to formulate all these ILPs by having a number of variables that is upper-bounded by a function of $k$. Finally we use Lenstra's Theorem~\cite{Lenstra1983Integer} to solve each of these ILPs in FPT time. At the end, our initial instance is a \textsc{Yes}-instance if and only if at least one of these ILPs is feasible.

@@

\subparagraph{Related work.} Graph realization problems on static graphs have been studied since the 1960s. We provide an overview of the literature in the introduction.

To the best of our knowledge, we are the first to consider graph realization problems in the temporal setting. However, many other connectivity-related problems have been studied in the temporal setting~\cite{Mertzios-transitivity21,Akrida-explorer-21,enright2021deleting,MolterRZ21,klobas2023interference,deligkas2022optimizing,erlebach2021temporal,Flu+19a,Zsc+19,CasteigtsCS22,FuchsleMNR22}, most of which are much more complex and computationally harder than their non-temporal counterparts, and some of which do not even have a non-temporal counterpart.

There are two problem settings that share similarities with our setting, which we discuss in more detail in the following.

Several problems have been studied where the goal is to assign labels to (sets of) edges of a given static graph in order to achieve certain connectivity-related properties~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity,enright2021assigning}. The main difference to our problem setting is that in the mentioned works, the input is a graph and the sought labeling is not periodic. Furthermore, the investigated properties are temporal connectivity among all vertices~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity}, temporal connectivity among a subset of vertices~\cite{KlobasMMS22}, or reducing reachability among the vertices~\cite{enright2021assigning}. In all cases the duration of the temporal connections is not of interest.

The only work we are aware of that studies problems on periodic temporal graphs is work on the temporal cops and robber game~\cite{Morawietz021,MorawietzRW20,ErlebachS20}. The question here is (in its basic case) whether a certain number of cops are able to catch a robber, if everybody can move along one edge at every time step and the robber is considered caught if it occupies the same vertex as a cop at the same time.

@@

\begin{figure}[t]
\noindent\makebox[\textwidth]{
\centering
%\begin{center}
\begin{tikzpicture}[line width=1pt, scale=.65, yscale=.6]
\scriptsize
\node (A) at (14,0) {};

@@

	\draw[edge, line width=1.5pt] (X1) --node[] {} (VS);
	\draw[edge] (X2) --node[] {} (VS);
	\draw[edge] (X3) --node[] {} (VS);
	\draw[edge] (XM) --node[] {} (VS);

    \draw[edge] (Y1) --node[] {} (VS);
	\draw[edge] (Y2) --node[] {} (VS);
	\draw[edge, line width=1.5pt] (Y3) --node[] {} (VS);
	\draw[edge] (YM) --node[] {} (VS);

    \draw[edge] (Y1) --node[] {} (U11);
	\draw[edge] (Y2) --node[] {} (U12);
	\draw[edge, line width=1.5pt] (Y3) --node[] {} (U13);
	\draw[edge] (YM) --node[] {} (U1M);

    \draw[edge] (U21) --node[] {} (U11);
	\draw[edge] (U22) --node[] {} (U12);
	\draw[edge, line width=1.5pt] (U23) --node[] {} (U13);
	\draw[edge] (U2M) --node[] {} (U1M);

    \draw[edge] (UN1) --node[] {} (UN11);
	\draw[edge] (UN2) --node[] {} (UN12);
	\draw[edge, line width=1.5pt] (UN3) --node[] {} (UN13);
	\draw[edge] (UNM) --node[] {} (UN1M);

    \draw[edge] (U2N1) --node[] {} (VSS);
	\draw[edge] (U2N2) --node[] {} (VSS);
	\draw[edge, line width=1.5pt] (U2N3) --node[] {} (VSS);
	\draw[edge] (U2NM) --node[] {} (VSS);

@@

	\draw[edge] (VII1) --node[] {} (UI1);
	\draw[edge] (UI1) --node[] {} (UI2);
\end{tikzpicture}
    %\end{center}
    }
    \caption{Illustration of part of the underlying graph $G$. Edges incident with vertices of connector gadgets are omitted. Gray vertices form a feedback vertex set.
    The double line connections, between a vertex $v_{i-1}^j$ in the verification gadget, and $u_1^3$ in the edge selection gadget, 
    and, between a vertex $u_2^3$ in the edge selection gadget, and $v_{i}^j$ in the verification gadget,
    consist of $5n$ vertices $a_1^{j,i,3},a_2^{j,i,3},\dots,a_{5n}^{j,i,3}$ 
    and $b_1^{j,i,3},b_2^{j,i,3},\dots,b_{5n}^{j,i,3}$, respectively.
    }\label{fig:hardness1}
\end{figure}

@@

Due to space constraints, we defer the description of the distance matrix $D$ and a formal correctness proof of the reduction to the Appendix.

\emph{Infinity gadget.} Finally, we show how to get rid of the infinity entries in $D$ and how to allow a finite $\Delta$. To this end, we introduce the \emph{infinity gadget}. We add four vertices $z_1, z_2, z_3, z_4$ to the graph and we set $\Delta=n^{11}$. Let $V$ denote the set of all remaining vertices. We set the following durations.
\begin{itemize}
    \item For all $v\in V$ we set $d(z_1,v)=2$, $d(z_2,v)=d(v,z_2)=1$, $d(z_3,v)=d(v,z_3)=1$, and $d(z_4,v)=2$.
    Furthermore, we set $d(v,z_1)=n^{11}$ and $d(v,z_4)=n^{10}-1$.
    \item We set $d(z_1,z_2)=d(z_2,z_1)=1$, $d(z_2,z_3)=d(z_3,z_2)=1$, and $d(z_3,z_4)=d(z_4,z_3)=1$.
    \item We set $d(z_1,z_3)=3$, $d(z_3,z_1)=n^{11}-1$, $d(z_2,z_4)=n^{10}-2$, and $d(z_4,z_2)=n^{11}-n^{10}+4$.
    \item We set $d(z_1,z_4)=n^{10}$ and $d(z_4,z_1)=2n^{11}-n^{10}+2$.
    \item For every pair of vertices $v,v'\in V$ where previously the duration of a fastest path from $v$ to $v'$ was specified to be infinite, we set $d(v,v')=n^{10}$.
\end{itemize}
Now we analyse which implications we get for the labels on the newly introduced edges.
Assume $\lambda(\{z_1,z_2\})=t$, then we get the following.
%\begin{itemize}
    %\item 
    For all $v\in V$ we have that $d(z_1,v)=2$ and hence we get that $\lambda(\{z_2,v\})=t+1$. 
    %\item 
    Since $d(z_1,z_4)=n^{10}$, we have that $\lambda(z_3,z_4)=t+n^{10}-1$.
    From this follows that for all $v\in V$, since $d(z_4,v)=2$, that $\lambda(\{z_3,v\})=t+n^{10}$.
    Finally, since $d(z_1,z_3)=3$, we have that $\lambda(\{z_2,z_3\})=t+2$.
%\end{itemize}
For an illustration see \cref{fig:hardness2}. It is easy to check that all duration requirements between vertex pairs in $\{z_1,z_2,z_3,z_4\}$ are met and that all duration requirements between each vertex $v\in V$ and each vertex in  $\{z_1,z_2,z_3,z_4\}$ are met. Furthermore, it is easy to check that the gadget increases the feedback vertex set by two ($z_2$ and $z_3$ need to be added).

@@

Lastly, consider two vertices $v,v'\in V$. Note that before the addition of the infinity gadget, by construction of $G$ we have that $d(v,v')\le n^9+2$ or $d(v,v')=\infty$. Furthermore, if $D$ is a \textsc{Yes}-instance, we have shown in the correctness proof of the reduction that the difference between the smallest label and the largest label is at most $n^9+1$. This implies that for a vertex pair $v,v'\in V$ with $d(v,v')=\infty$ we have in the periodic case with $\Delta=n^{11}$, that $d(v,v')\ge n^{11}-n^9>n^{10}$. Which means, after adding the vertices and edges of the infinity gadget, we indeed have that $d(v,v')=n^{10}$. For all vertex pairs $v,v'$ where in the original construction we have $d(v,v')\neq\infty$, we can also see that adding the infinity gadget and setting $\Delta=n^{11}$ does not change the duration of a fastest path from $v$ to $v'$, since all newly added temporal paths have duration at least $n^{10}$. We can conclude that the originally constructed instance $D$ is a \textsc{Yes}-instance if and only if it remains a \textsc{Yes}-instance after adding the infinity gadget and setting $\Delta=n^{11}$.
\end{proof}

@@

\section{Algorithms for \deltaExact}
In this section we provide several algorithms for \deltaExact. By \cref{thm:NPhardness} we have that \deltaExact\ is NP-hard in general, hence we start by identifying restricted cases where we can solve the problem in polynomial time.

We first show in \cref{sec:treealgo} that in the case when the underlying graph $G$ of an instance $(D, \Delta)$ of \deltaExact\ is a tree, then we can determine desired $\Delta$-periodic labeling $\lambda$ of $G$ in polynomial time. In \cref{sec:FPT} we generalize this result. We show that \deltaExact\ is fixed-parameter tractable when parameterized by the feedback edge number of the underlying graph. Note that our parameterized hardness result (\cref{thm:W1wrtFVS}) implies that we presumably cannot replace the feedback edge number with the smaller parameter feedback vertex number, or any other parameter that is smaller than the feedback vertex number, such as e.g.\ the treewidth.
In the Appendix we give also a dedicated algorithm to solve \deltaExact\ in polynomial time if the underlying graph is a cycle. Note that this is already implied by our FPT result, however the direct algorithm we provide is much simpler for this case.

\subsection{Polynomial-time algorithm for trees}\label{sec:treealgo}
We first start with defining certain notions, that will be of use when solving the problem.

>>>>

The \emph{graph realization} problem with respect to a graph property $\mathcal{P}$ is to find a graph that satisfies property $\mathcal{P}$, or to decide that no such graph exists. 
The motivation for graph realization problems stems both from ``verification'' and from network design applications in engineering. 
In \emph{verification} applications, given the outcomes of some experimental measurements (resp.~some computations) on a network, 
the aim is to (re)construct an input network which complies with them. 
If such a reconstruction is not possible, this proves that the measurements are incorrect or implausible (resp.~that the algorithm which made the computations is incorrectly implemented). 
One example of a graph realization (or reconstruction) problem is the recognition of probe interval graphs, in the context of the physical mapping of DNA, see~\cite{McMorris98,McConnellS02} and~\cite[Chapter 4]{GolumbicTrenk04}.
In \emph{network design} applications, the goal is to design network topologies having a desired property~\cite{augustine2022distributed,grotschel1995design}.
Analyzing the computational complexity of the graph realization problems for various natural and fundamental graph properties $\mathcal{P}$ requires a deep understanding of these properties.
Among the most studied such parameters for graph realization 
are constraints on the distances between vertices~\cite{barNoy2022GraphRealization,barNoy2021composed,hakimi1965distance,chung2001distance,bixby1988almost,culberson1989fast}, 
on the vertex degrees~\cite{GolovachM17,gomory1961multi,hakimi1962realizability,Bar-NoyCPR20,erdos1960graphs}, 
%%%%barNoy2021selected,cai2000solution
on the eccentricities~\cite{barNoy2020efficiently,hell2009linear,behzad1976eccentric,lesniak1975eccentric}, and on connectivity~\cite{fulkerson1960zero,frank1992augmenting,chen1966realization,frank1994connectivity,frank1970connectivity,gomory1961multi}, among others.

In the simplest version of a graph realization problem with respect to vertex distances, 
we are given a symmetric $n \times n$ matrix $D$ and we are looking for an $n$-vertex undirected and unweighted graph $G$ such that $D_{i,j}$ equals the distance between vertices $v_i$ and $v_j$ in~$G$. This problem can be trivially solved in polynomial time in two steps~\cite{hakimi1965distance}: First, we build the graph $G=(V,E)$ such that $v_i v_j \in E$ if and only if $D_{i,j}=1$. Second, from this graph $G$ we compute the matrix $D_G$ which captures the shortest distances for all pairs of vertices. If $D_G = D$ then $G$ is the desired graph, otherwise there is no graph having $D$ as its distance matrix. 
Non-trivial variations of this problem have been extensively studied, such as for weighted graphs~\cite{hakimi1965distance,Patrinos-Hakimi-72}, as well as for cases where the realizing graph has to belong to a specific graph family~\cite{hakimi1965distance, barNoy2021composed}. Other variations of the problem include the cases where every entry of the input matrix $D$ may contain a range of consecutive permissible values~\cite{barNoy2021composed,Rubei16,Tamura93}, or even an arbitrary set of acceptable values~\cite{barNoy2022GraphRealization} for the distance between the corresponding two vertices.

@@

In this paper we consider \emph{periodic} temporal graphs, i.e.~temporal graphs in which the temporal availability of each edge of the underlying graph is periodic. 
Many natural and technological systems exhibit a periodic temporal behavior. For example, in railway networks an edge is present at a time step $t$ if and
only if a train is scheduled to run on the respective rail segment at time $t$~\cite{Arrighi2023Multi}. 
Similarly, a satellite, which makes pre-determined periodic movements, can establish a communication link (i.e.~a temporal edge) with another satellite whenever they are sufficiently close to each other; the existence of these communication links is also periodic. 
In a railway (resp.~satellite) network, a fastest temporal path from $u$ to $v$ represents the fastest railway connection between two stations 
(resp.~the quickest communication delay between two moving satellites). 
Furthermore, periodicity appears also in (the otherwise quite complex) social networks which describe the dynamics of people meeting~\cite{snapnets,sapiezynski2015tracking}, as every person individually follows mostly a daily routine~\cite{Arrighi2023Multi}.

Although periodic temporal graphs have already been studied 
(see~\cite[Class 8]{casteigts2012time} and~\cite{Arrighi2023Multi,ErlebachS20,morawietz2021timecop,morawietz2020timecop}), 
we make here the first attempt to understand the complexity of a graph realization problem in the context of temporal graphs. 
Therefore, we focus in this paper on the most fundamental case, where all edges have the same period $\Delta$ 
(while in the more general case, each edge $e$ in the underlying graph has a period $\Delta_e$).
As it turns out, the periodic temporal graph realization problem with respect to a given $n \times n$ matrix $D$ of the fastest duration times has a very different computational complexity behavior than the classic graph realization problem with respect to shortest path distances in static graphs.

%In an arbitrary (non-periodic) temporal graph $(G,\lambda)$ there can be pairs of vertices $u,v$ where there is no temporal path at all from $u$ to $v$, even if the underlying graph $G$ is undirected and connected.

%In order to focus on the smallest duration of temporal paths, rather than on temporal reachability between vertices, we consider in this paper only \emph{periodic} temporal graphs, i.e.~on temporal graphs in which the availability times of every edge are periodic (all edges having the same period $\Delta$)~\cite{casteigts2012time}.

%Apart from the theoretical interest of periodic temporal graphs, which have not been well studied and understood in the literature, they capture naturally many application areas too, such as transportation or satellite networks, where entities move according to repeating schedules or time tables. For example, a fastest temporal path from vertex $u$ to vertex $v$ in the temporal graph defined by a transportation (resp.~satellite) network represents the shortest travel time (resp.~the shortest communication delay) from $u$ to $v$.

@@

\subparagraph{Our contribution.}
%In this work, 
We initiate the study of naturally motivated graph realization problems in the temporal setting. 
Our target is not to model unreliable communication, but instead to \emph{verify} that particular measurements regarding fastest temporal paths in a periodic temporal graph are plausible (i.e.~``realizable''). 
To this end, we introduce and investigate the following problem, capturing the setting described above:

@@

We focus on exact algorithms. We start with showing NP-hardness of the problem (\cref{thm:NPhardness}), even if $\Delta$ is a small constant. To establish a baseline for tractability, we show that \deltaExact\ is polynomial-time solvable if the underlying graph is a tree (\cref{thm:deltaExact-PolyTimeTrees}).

@@

We obtain the following results. We show that \deltaExact\ is W[1]-hard when parameterized by the feedback vertex number of the underlying graph (\cref{thm:W1wrtFVS}). 
To this end, we first give a reduction from \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval} to a variant of \deltaExact\ where the period $\Delta$ is infinite, that is, when the labeling is non-periodic. We use a special gadget (the ``infinity'' gadget) which allows us to transfer the result to a finite period $\Delta$. The latter construction is independent from the particular reduction we use, and can hence be treated as a reduction from the non-periodic to the periodic setting.
Note that our parameterized hardness result rule out fixed-parameter tractability for several popular graph parameters such as treewidth and arboricity.

We complement this result by showing that the problem is fixed-parameter tractable (FPT) with respect to the feedback edge number $k$ of the underlying graph (\cref{thm:FPTwrtFES}), a parameter that is larger than the feedback vertex number. 
Our FPT algorithm works as follows on a high level. 
First we distinguish $O(k)$ vertices which we call ``interesting vertices''. 
Then, we guess the fastest temporal paths for each pair of these interesting vertices; as we prove, the number of choices we have for all these guesses is upper bounded by a function of~$k$. 
Then we also make several further guesses (again using a bounded number of choices), which altogether leads us to specify a small (i.e.~bounded by a function of $k$) number of different configurations for the fastest paths between \emph{all pairs} of vertices. For each of these configurations, we must then make sure that the labels of our solution will not allow any other temporal path from a vertex $v_i$ to a vertex $v_j$ have a \emph{strictly smaller} duration than~$D_{i,j}$.
This naturally leads us to build one Integer Linear Program (ILP) for each of these configurations. We manage to formulate all these ILPs by having a number of variables that is upper-bounded by a function of $k$. Finally we use Lenstra's Theorem~\cite{Lenstra1983Integer} to solve each of these ILPs in FPT time. At the end, our initial instance is a \textsc{Yes}-instance if and only if at least one of these ILPs is feasible.

@@

\subparagraph{Related work.} Graph realization problems on static graphs have been studied since the 1960s. We provide an overview of the literature in the introduction. 
To the best of our knowledge, we are the first to consider graph realization problems in the temporal setting. However, many other connectivity-related problems have been studied in the temporal setting~\cite{Mertzios-transitivity21,Akrida-explorer-21,enright2021deleting,MolterRZ21,klobas2023interference,deligkas2022optimizing,erlebach2021temporal,Flu+19a,Zsc+19,CasteigtsCS22,FuchsleMNR22}, most of which are much more complex and computationally harder than their non-temporal counterparts, and some of which do not even have a non-temporal counterpart.

There are some problem settings that share similarities with ours, which we discuss now in more detail.

Several problems have been studied where the goal is to assign labels to (sets of) edges of a given static graph in order to achieve certain connectivity-related properties~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity,enright2021assigning}. The main difference to our problem setting is that in the mentioned works, the input is a graph and the sought labeling is not periodic. Furthermore, the investigated properties are temporal connectivity among all vertices~\cite{KlobasMMS22,MertziosMS19,akrida2017complexity}, temporal connectivity among a subset of vertices~\cite{KlobasMMS22}, or reducing reachability among the vertices~\cite{enright2021assigning}. In all these cases, the duration of the temporal paths has not been considered.

%work on the temporal cops and robber game~\cite{morawietz2021timecop,morawietz2020timecop,ErlebachS20}. The question here is (in its basic case) whether a certain number of cops are able to catch a robber, if everybody can move along one edge at every time step and the robber is considered caught if it occupies the same vertex as a cop at the same time. Several other works on periodic temporal graphs were mentioned in the introduction.

Finally, there are many models for dynamic networks in the context of distributed computing~\cite{Kuhn2011Dynamic}. 
These models have some similarity to temporal graphs, in the sense that in both cases the edges appear and disappear over time. 
However, there are notable differences. For example, one important assumption in the distributed setting 
can be that the edge changes are adversarial or random (while obeying some constraints such as connectivity), 
and therefore they are not necessarily known in advance~\cite{Kuhn2011Dynamic}.

@@

\begin{figure}[t!]
\noindent\makebox[\textwidth]{
\centering
%\begin{center}
\begin{tikzpicture}[line width=1pt, scale=.65, yscale=.7]
\scriptsize
\node (A) at (14,0) {};

@@

	\draw[edge] (X1) --node[] {} (VS);
	\draw[edge] (X2) --node[] {} (VS);
	\draw[edge] (X3) --node[] {} (VS);
	\draw[edge] (XM) --node[] {} (VS);

    \draw[edge] (Y1) --node[] {} (VS);
	\draw[edge] (Y2) --node[] {} (VS);
	\draw[edge] (Y3) --node[] {} (VS);
	\draw[edge] (YM) --node[] {} (VS);

    \draw[edge] (Y1) --node[] {} (U11);
	\draw[edge] (Y2) --node[] {} (U12);
	\draw[edge] (Y3) --node[] {} (U13);
	\draw[edge] (YM) --node[] {} (U1M);

    \draw[edge] (U21) --node[] {} (U11);
	\draw[edge] (U22) --node[] {} (U12);
	\draw[edge] (U23) --node[] {} (U13);
	\draw[edge] (U2M) --node[] {} (U1M);

    \draw[edge] (UN1) --node[] {} (UN11);
	\draw[edge] (UN2) --node[] {} (UN12);
	\draw[edge] (UN3) --node[] {} (UN13);
	\draw[edge] (UNM) --node[] {} (UN1M);

    \draw[edge] (U2N1) --node[] {} (VSS);
	\draw[edge] (U2N2) --node[] {} (VSS);
	\draw[edge] (U2N3) --node[] {} (VSS);
	\draw[edge] (U2NM) --node[] {} (VSS);

@@

	\draw[edge] (VII1) --node[] {} (UI1);
	\draw[edge] (UI1) --node[] {} (UI2);
\end{tikzpicture}
    %\end{center}
    }
    \caption{Illustration of part of the underlying graph $G$. Edges incident with vertices $\hat{v}_1,\hat{v}_2$ of connector gadgets are omitted. Gray vertices form a feedback vertex set.
    The double line connections, between a vertex $v_{i-1}^j$ in the verification gadget, and $u_1^3$ in the edge selection gadget, 
    and, between a vertex $u_2^3$ in the edge selection gadget, and $v_{i}^j$ in the verification gadget,
    consist of $5n$ vertices $a_1^{j,i,3},a_2^{j,i,3},\dots,a_{5n}^{j,i,3}$ 
    and $b_1^{j,i,3},b_2^{j,i,3},\dots,b_{5n}^{j,i,3}$, respectively.
    }\label{fig:hardness1}
\end{figure}

@@

Due to space constraints, we defer the description of the distance matrix $D$ and a formal correctness proof of the reduction to the Appendix.
%
\begin{comment}
\emph{Infinity gadget.} Finally, we show how to get rid of the infinity entries in $D$ and how to allow a finite $\Delta$. To this end, we introduce the \emph{infinity gadget}. We add four vertices $z_1, z_2, z_3, z_4$ to the graph and we set $\Delta=n^{11}$. Let $V$ denote the set of all remaining vertices. We set the following durations.
\begin{itemize}
    \item For all $v\in V$ we set $d(z_1,v)=2$, $d(z_2,v)=d(v,z_2)=1$, $d(z_3,v)=d(v,z_3)=1$, and $d(z_4,v)=2$.
    Furthermore, we set $d(v,z_1)=n^{11}$ and $d(v,z_4)=n^{10}-1$.
    \item We set $d(z_1,z_2)=d(z_2,z_1)=1$, $d(z_2,z_3)=d(z_3,z_2)=1$, and $d(z_3,z_4)=d(z_4,z_3)=1$.
    \item We set $d(z_1,z_3)=3$, $d(z_3,z_1)=n^{11}-1$, $d(z_2,z_4)=n^{10}-2$, and $d(z_4,z_2)=n^{11}-n^{10}+4$.
    \item We set $d(z_1,z_4)=n^{10}$ and $d(z_4,z_1)=2n^{11}-n^{10}+2$.
    \item For every pair of vertices $v,v'\in V$ where previously the duration of a fastest path from $v$ to $v'$ was specified to be infinite, we set $d(v,v')=n^{10}$.
\end{itemize}
Now we analyse which implications we get for the labels on the newly introduced edges.
Assume $\lambda(\{z_1,z_2\})=t$, then we get the following.
%\begin{itemize}
    %\item 
    For all $v\in V$ we have that $d(z_1,v)=2$ and hence we get that $\lambda(\{z_2,v\})=t+1$. 
    %\item 
    Since $d(z_1,z_4)=n^{10}$, we have that $\lambda(z_3,z_4)=t+n^{10}-1$.
    From this follows that for all $v\in V$, since $d(z_4,v)=2$, that $\lambda(\{z_3,v\})=t+n^{10}$.
    Finally, since $d(z_1,z_3)=3$, we have that $\lambda(\{z_2,z_3\})=t+2$.
%\end{itemize}
For an illustration see \cref{fig:hardness2}. It is easy to check that all duration requirements between vertex pairs in $\{z_1,z_2,z_3,z_4\}$ are met and that all duration requirements between each vertex $v\in V$ and each vertex in  $\{z_1,z_2,z_3,z_4\}$ are met. Furthermore, it is easy to check that the gadget increases the feedback vertex set by two ($z_2$ and $z_3$ need to be added).

@@

Lastly, consider two vertices $v,v'\in V$. Note that before the addition of the infinity gadget, by construction of $G$ we have that $d(v,v')\le n^9+2$ or $d(v,v')=\infty$. Furthermore, if $D$ is a \textsc{Yes}-instance, we have shown in the correctness proof of the reduction that the difference between the smallest label and the largest label is at most $n^9+1$. This implies that for a vertex pair $v,v'\in V$ with $d(v,v')=\infty$ we have in the periodic case with $\Delta=n^{11}$, that $d(v,v')\ge n^{11}-n^9>n^{10}$. Which means, after adding the vertices and edges of the infinity gadget, we indeed have that $d(v,v')=n^{10}$. For all vertex pairs $v,v'$ where in the original construction we have $d(v,v')\neq\infty$, we can also see that adding the infinity gadget and setting $\Delta=n^{11}$ does not change the duration of a fastest path from $v$ to $v'$, since all newly added temporal paths have duration at least $n^{10}$. We can conclude that the originally constructed instance $D$ is a \textsc{Yes}-instance if and only if it remains a \textsc{Yes}-instance after adding the infinity gadget and setting $\Delta=n^{11}$.
\end{comment}
\end{proof}

@@

\section{Algorithms for \deltaExact}
In this section we provide several algorithms for \deltaExact. By \cref{thm:NPhardness} we have that \deltaExact\ is NP-hard in general, hence we start by identifying restricted cases where we can solve the problem in polynomial time.
%
The first case, when the underlying graph $G$ of an instance $(D, \Delta)$ of \deltaExact\ is a tree, we get the following result.

\begin{theorem}[$\star$]\label{thm:deltaExact-PolyTimeTrees}
    \deltaExact\ can be solved in polynomial-time on trees.
\end{theorem}

%We first show %in \cref{sec:treealgo} 
%that in the case when the underlying graph $G$ of an instance $(D, \Delta)$ of \deltaExact\ is a tree, then we can determine desired $\Delta$-periodic labeling $\lambda$ of $G$ in polynomial time. 
In the case when the underlying graph is unrestricted, we show that \deltaExact\ is fixed-parameter tractable when parameterized by the feedback edge number of the underlying graph. Note that our parameterized hardness result (\cref{thm:W1wrtFVS}) implies that we presumably cannot replace the feedback edge number with the smaller parameter feedback vertex number, or any other parameter that is smaller than the feedback vertex number, such as e.g.\ the treewidth.
In the Appendix we give also a dedicated algorithm to solve \deltaExact\ in polynomial time if the underlying graph is a cycle. Note that this is already implied by our FPT result, however the direct algorithm we provide is much simpler for this case.

% \subsection{Polynomial-time algorithm for trees}\label{sec:treealgo}
% We first start with defining certain notions, that will be of use when solving the problem.
